Characterizing Water Composition with an Autonomous Robotic Team Employing Comprehensive In-Situ Sensing, Hyperspectral Imaging, Machine Learning, and Conformal Prediction

### 2024-02-19 paper/paper.tex
Updates from Overleaf

The USV employed in the robot team is a Maritime Robotics Otter which features a modular design and 20 hour battery life. The USV is equipped with an in situ sensing payload consisting of a combination of Eureka Manta-40 multiprobes including fluorometers, ion selective electrodes, and other physical sensors that together enable the collection of comprehensive near-surface measurements including Colored Dissolved Organic Matter (CDOM), Crude Oil, Blue-Green Algae (phycoerythrin and phycocyanin), Chlorophyll A, $\mathrm{Na^+}$, $\mathrm{Ca^{2+}}$, $\mathrm{Cl^-}$, Temperature, Conductivity, and many others. The full list of measurements utilized in this study is given in Table~\ref{tab:sensors} and is categorized into four distinct types: physical measurements, ion measurements, biochemical measurements, and chemical measurements. Additionally, the USV is equipped with an ultrasonic weather monitoring sensor for measuring air speed and direction, as well as a BioSonics MX Aquatic Habitat Echosounder sonar, which are not explored in this study.

@@

\begin{table}[H] 
\caption{In-situ reference sensors modelled in this study.\label{tab:sensors}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCCC}
\toprule
\textbf{Sensor}	& \textbf{Units} & Resolution & \textbf{Sensor Type} & \textbf{Target Category}\\
\midrule
Temperature		                    & $^{\circ}$C        & 0.01  & Thermistor                           & Physical \\
Conductivity                        & $\mu$S/cm          & 0.01  & Four-Electrode Graphite Sensor       & Physical  \\
pH                                  & logarithmic (0-14) & 0.01  & Flowing-junction Reference Electrode & Physical \\
Turbidity                           & FNU                & 0.01  & Ion-Selective Electrode              & Physical \\
$\mathrm{Ca^{2+}}$                  & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Cl^-}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Na^+}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
Blue-Green Algae (phycoerythrin)    & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Blue-Green Algae (phycocyanin)      & ppb                & 0.01  & Fluorometer                          & Biochemical \\
CDOM                                & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Chlorophyll A                       & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Optical Brighteners                 & ppb                & 0.01  & Fluorometer                          & Chemical \\
Crude Oil                           & ppb                & 0.01  & Fluorometer                          & Chemical\\
\bottomrule
\end{tabularx}
% \noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
\end{table}

@@

For each of the 13 target variables listed in Table~\ref{tab:sensors}, the data was randomly partitioned into a 90:10 training/testing split. Using the MLJ machine learning framework, a variety of machine learning models compatible with the regression task were identified \cite{MLJ1}. This selection of models included linear regression (and it's variations), Gaussian process regression, shallow neural networks, and decision tree-based models. Out of this set, the Random Forest Regressor (RFR) model from \cite{decision-trees} was selected for its relatively quick training and inference times. Random forests are an ensembling technique based on bagged predictions of individual decision tree regressors trained using the CART algorithm with each individual tree trained on a random subset of features \textit{and} a random subset of training records \cite{decision-trees, random-forest}.

@@

Beyond the potential applications of this robot team for direct water quality monitoring, the combination of reference instruments together with easily co-located hyperspectral imagery provides fertile ground for the development of new spectral indices targeted towards specific water quality variables. In this paper, we have shown how permutation importance ranking for trained machine learning models enables a straightforward interpretation of the relative value of each reflectance bin to the final model predictions. In future work, we plan to utilize this information to construct new combinations of spectral bands which can be readily transferred to remote sensing platforms equipped with hyperspectral imagers such as the German EnMAP mission.

>>>>

\begin{figure}
\includegraphics[width=\columnwidth]{paper/figures/materials-and-methods/annotated-boat.pdf}
\caption{ Configuration of the USV: (\textbf{a}) Frontal view of the USV showing the Eureka Manta+40 Multiprobes mounted on the underside of the boat. (\textbf{b}) The USV deployed in the water.\label{fig:boat-components}}
\end{figure}

The USV employed in the robot team is a Maritime Robotics Otter equipped with an in situ sensing payload consisting of a combination of Eureka Manta+40 multiprobes. These sensors include fluorometers, ion selective electrodes, and other physical sensors and are mounted on the underside of boat as illustrated in Figure~\ref{fig:boat-components}. Together this sensor array enables the collection of comprehensive near-surface measurements including Colored Dissolved Organic Matter (CDOM), Crude Oil, Blue-Green Algae (phycoerythrin and phycocyanin), Chlorophyll A, $\mathrm{Na^+}$, $\mathrm{Ca^{2+}}$, $\mathrm{Cl^-}$, Temperature, Conductivity, and many others. The full list of measurements utilized in this study is outlined in Table~\ref{tab:sensors} and is categorized into four distinct types: physical measurements, ion measurements, biochemical measurements, and chemical measurements. Additionally, the USV is equipped with an ultrasonic weather monitoring sensor for measuring air speed and direction, as well as a BioSonics MX Aquatic Habitat Echosounder sonar, which are not explored in this study.

@@

\begin{table}[H] 
\caption{In-situ reference sensors utilized in this study.\label{tab:sensors}}
\newcolumntype{C}{>{\centering\arraybackslash}X}
\begin{tabularx}{\textwidth}{CCCCC}
\toprule
\textbf{Sensor}	& \textbf{Units} & Resolution & \textbf{Sensor Type} & \textbf{Target Category}\\
\midrule
Temperature		                    & $^{\circ}$C        & 0.01  & Thermistor                           & Physical \\
Conductivity                        & $\mu$S/cm          & 0.01  & Four-Electrode Graphite Sensor       & Physical  \\
pH                                  & logarithmic (0-14) & 0.01  & Flowing-junction Reference Electrode & Physical \\
Turbidity                           & FNU                & 0.01  & Ion-Selective Electrode              & Physical \\
$\mathrm{Ca^{2+}}$                  & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Cl^-}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
$\mathrm{Na^+}$                     & mg/l               & 0.1   & Ion-Selective Electrode              & Ions \\
Blue-Green Algae (phycoerythrin)    & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Blue-Green Algae (phycocyanin)      & ppb                & 0.01  & Fluorometer                          & Biochemical \\
CDOM                                & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Chlorophyll A                       & ppb                & 0.01  & Fluorometer                          & Biochemical \\
Optical Brighteners                 & ppb                & 0.01  & Fluorometer                          & Chemical \\
Crude Oil                           & ppb                & 0.01  & Fluorometer                          & Chemical\\
\bottomrule
\end{tabularx}
% \noindent{\footnotesize{\textsuperscript{1} Tables may have a footer.}}
\end{table}

@@

For each of the 13 target variables listed in Table~\ref{tab:sensors}, the data was randomly partitioned into a 90:10 training/testing split. Using Julia's MLJ (Machine Learning Julia) package, a variety of machine learning models compatible with the regression task were identified \cite{MLJ1}. This selection of models included linear regression (and it's variations), Gaussian process regression, shallow neural networks, and decision tree-based models. Out of this set, the Random Forest Regressor (RFR) model from \cite{decision-trees} was selected for use in this study due to its relatively quick training and inference times. Random forests are an ensembling technique based on bagged predictions of individual decision tree regressors trained using the classification and regression trees (CART) algorithm. Each individual tree in a RFR is trained on a random subset of features \textit{and} a random subset of training records \cite{decision-trees, random-forest}.

@@

Beyond the potential applications of this robot team for direct water quality monitoring, the combination of reference instruments together with easily co-located hyperspectral imagery provides fertile ground for the development of new spectral indices targeted towards specific water quality variables. In this paper, we have shown how permutation importance ranking for trained machine learning models enables a straightforward interpretation of the relative value of each reflectance bin to the final model predictions. In future work, we plan to utilize this information to construct new combinations of spectral bands which can be readily extended for use remote sensing platforms equipped with hyperspectral imagers. \textcolor{red}{The recently launched Environmental Mapping and Analysis Program (EnMAP) is one such example which includes over 91 spectral bands in the VNIR that overlap with our hyperspectral imager \cite{Storch2023TheEI}.}
